---
title: "Lab 8"
author: "Kristina Arevalo"
date: '`r format(Sys.time(), "%a %b %d %Y")`'
output: 
  pdf_document:
    latex_engine: xelatex
    toc: yes 
    toc_depth: 2
monofont: Times New Roman
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE, message = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
library(patchwork)
```

\newpage

# Problem 1A
Consider a 2x2 design. Assume the DV is measured from a normal distribution with mean 0, and standard deviation 1. Assume that the main effect of A causes a total shift of .5 standard deviations of the mean between the levels. Assume that level 1 of B is a control, where you expect to measure the standard effect of A. Assume that level 2 of B is an experimental factor intended to reduce the effect of A by .25 standard deviations.

Create a ggplot2 figure that depicts the expected results from this design (2 points)

```{r}
grand_mean <- 0
A <- c(0,.5)
B <- c(0,0)
AB <- c(0,0,0,-.25)

model_data <- tibble()
for(i in 1:length(A)){
  for(j in 1:length(B)){
    IVA <- i
    IVB <- j
    DV <- grand_mean+A[i]+B[j]+AB[(i-1)*length(B)+j]
    sc_GM <- grand_mean
    sc_A <- A[i]
    sc_B <- B[j]
    sc_AB <- AB[(i-1)*length(B)+j]
    row_entry <- tibble(IVA,IVB,DV,
                        sc_GM,sc_A,sc_B,sc_AB)
    model_data <- rbind(model_data,row_entry)
  }
}

knitr::kable(model_data)

bar_graph <- ggplot(model_data,
                    aes(y=DV,
                        x=as.factor(IVA),
                        fill=as.factor(IVB)))+
  geom_bar(stat='identity', position='dodge')

line_graph <- ggplot(model_data,
                     aes(y=DV,
                         x=IVA,
                         linetype=as.factor(IVB)))+
  geom_line()+
  geom_point()

(bar_graph/line_graph)
```


# Problem 1B
Conduct simulation-based power analyses to answer the questions.

How many subjects are needed to detect the main effect of A with power = .8? (2 points)

```{r}
N <- 57

A_pvalue <- c()
B_pvalue <- c()
AB_pvalue <- c()
for(i in 1:1000){
  IVA <- rep(rep(c("1","2"), each=2),N)
  IVB <- rep(rep(c("1","2"), 2),N)
  DV <- c(replicate(N,c(rnorm(1,0,1), 
                        rnorm(1,0,1), 
                        rnorm(1,.5,1),
                        rnorm(1,.25,1) 
          )))
  sim_df <- data.frame(IVA,IVB,DV)

  aov_results <- summary(aov(DV~IVA*IVB, sim_df))
  A_pvalue[i]<-aov_results[[1]]$`Pr(>F)`[1]
  B_pvalue[i]<-aov_results[[1]]$`Pr(>F)`[2]
  AB_pvalue[i]<-aov_results[[1]]$`Pr(>F)`[3]
}

length(A_pvalue[A_pvalue<0.05])/1000

```

# Problem 1C
How many subjects are needed to detect the interaction effect with power = .8? (2 points)
```{r}
N <- 490

A_pvalue <- c()
B_pvalue <- c()
AB_pvalue <- c()
for(i in 1:1000){
  IVA <- rep(rep(c("1","2"), each=2),N)
  IVB <- rep(rep(c("1","2"), 2),N)
  DV <- c(replicate(N,c(rnorm(1,0,1), 
                        rnorm(1,0,1), 
                        rnorm(1,.5,1),
                        rnorm(1,.25,1) 
          )))
  sim_df <- data.frame(IVA,IVB,DV)

  aov_results <- summary(aov(DV~IVA*IVB, sim_df))
  A_pvalue[i]<-aov_results[[1]]$`Pr(>F)`[1]
  B_pvalue[i]<-aov_results[[1]]$`Pr(>F)`[2]
  AB_pvalue[i]<-aov_results[[1]]$`Pr(>F)`[3]
}

length(AB_pvalue[AB_pvalue<0.05])/1000

```


