---
title: "Lab 4"
author: "Kristina Arevalo"
date: '`r format(Sys.time(), "%a %b %d %Y")`'
output: 
  pdf_document:
    toc: yes 
    toc_depth: 2
monofont: Times New Roman
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE, message = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(papaja)
library(tibble)
```

\newpage
# Problem 1 
Consider the following example data for a between-subjects experiment with two groups, A and B:

```{r}
example_data <- tibble(Group = rep(c("A","B"), each = 5),
                       DV = c(2,4,3,5,4,7,6,5,6,7))
```
                       
Use R to conduct a t.test and ANOVA on this data. Then use R to prove that the results of both analyses are the same. For example, prove that the p-values are the same, and prove that the F-value and T-value are related. (3 points)

```{r}
t_test_output <- t.test(DV~Group, var.equal = TRUE, example_data)
```

```{r}
anova_output <- aov(DV~Group,example_data) %>% summary()
```

To prove that the p-values are the same: 
```{r}
t_test_output$p.value
anova_output[[1]]$`Pr(>F)`[1]
```

To show that t-value& F-value are related: 
```{r}
(t_test_output$statistic)^2
anova_output[[1]]$`F value`[1]
```

\newpage
# Problem 2
Look at the lab on ANOVA that I wrote for our undergraduate statistics OER lab manual https://crumplab.github.io/statisticsLab/lab-8-one-way-anova.html. That lab shows an example of obtaining data from a published paper in psych science where a one-factor ANOVA was used as a part of the analysis. Load the data, conduct the ANOVA, report a ggplot of the means, and use papaja to help you write a short results section reporting the ANOVA result. (3 points).

```{r}
library(data.table)
all_data <- read.csv("open_data/Jamesetal2015Experiment2.csv")

all_data$Condition <- as.factor(all_data$Condition)
levels(all_data$Condition) <- c("Control",
                                "Reactivation+Tetris", 
                                "Tetris_only",
                                "Reactivation_only")
```

## ANOVA
```{r}
anova_jamesetal<- aov(Days_One_to_Seven_Number_of_Intrusions~Condition, all_data) %>% summary()

anova_jamesetal
```

## Figure
```{r}
ggplot(all_data, aes(x= Condition, y= Days_One_to_Seven_Number_of_Intrusions, fill = Condition)) +
  geom_bar(stat= "summary", fun = "mean", position ="dodge")+
  geom_point()+
  theme_classic()
```


## Results
```{r}
apa_print(anova_jamesetal)$full_result$Condition
```

A one-factor between-subjects ANOVA was conducted with Intervention type as the independent variable. We found a main effect of intervention type, `r apa_print(anova_jamesetal)$full_result$Condition`

















