---
title: "Lab 2"
author: "Kristina Arevalo"
date: '`r format(Sys.time(), "%a %b %d %Y")`'
output: 
  pdf_document:
    toc: yes 
    toc_depth: 2
monofont: Times New Roman
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE, message = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
```

\newpage
# Problem 1 
The following code roughly reproduces figure 5.5 from the textbook (Abdi et al., 2009). Modify the ggplot code so that the graph looks as close as possible to the figure from textbook:
  change the x-axis title
  change the y-axis title
  make the x-axis and y axis ticks the same
  include different symbols to differentiate the lines


```{r}
slamecka_design <- tibble(number_of_learning_trials = rep(c(2,4,8), each=6),
                          number_of_IL = rep(rep(c(2,4,8), 2), 3),
                          subjects = 1:18,
                          recall = c(35,21,6,
                                     39,31,8,
                                     40,34,18,
                                     52,42,26,
                                     61,58,46,
                                     73,66,52
                          )
)

slamecka_design$number_of_learning_trials <- as.factor(slamecka_design$number_of_learning_trials)

ggplot(slamecka_design,aes(x=number_of_IL,
                           group = number_of_learning_trials,
                           y=recall,
                           shape = number_of_learning_trials))+
         geom_line(stat = "summary", fun = "mean")+
         geom_point(stat = "summary", fun = "mean",)+
         theme_classic()+
         labs(x = "Number of Interpolated Lists", y = "Number of Words Correct")+
         scale_y_continuous(breaks = c(20, 40, 60 ,80), limits = c(0,80))+
         scale_x_continuous(breaks= c(2,4,8))


```
I could not figure out how to label the lines instead of having a legend. There were so many ways to do it and none of them worked for me. I tried geom_dl (Directlabel) but it kept saying that function didn't exist and I tried geom_label but it just made boxes of the numbers around each point. If you know of any easier way, I would appreciate it! 

# Problem 2 

The Slamecka design was a 3x3 design. The practice variable had three levels (2, 4, and 8), and was assumed to increase memory recall with increasing amount of practice. The interpolated learning (IL or distraction) variable had three levels (0, 4, and 8), and was assumed to decrease memory recall with increasing amounts of distraction.

Imagine Slamecka had included a third independent variable that also had three levels. This new IV is expected to increase memory recall. For example, maybe the third IV is amount of reward given to subjects (0, 50 dollars, 1 million dollars) for completing the study. As the amount of expected reward is increased, subjects are more motivated so they remember more words.

Use R to create a dataframe to represent predictions for this new 3x3x3 design. Then use ggplot and facet wrapping to show predicted means for each condition. As a hint, I have provided a drawing and I expect your ggplot graph will look something like this drawing.

```{r}
slamecka_3x3 <- tibble(money_reward = rep(c("A:0","B: 50", "C:1000000"), each = 9),
                       amount_of_practice = rep(rep(c(2,4,8), each = 3),3),
                       distraction = as.factor(rep(c(0,4,8), 9)),
                       recall = c(5,3,1,
                                   6,4,2,
                                   8,6,4,
                                   10,8,6,
                                   12,10,8,
                                   14,12,10,
                                   16,14,12,
                                   18,16,14,
                                   20,18,16
                       )
                       
)
ggplot(slamecka_3x3,aes(x=amount_of_practice,
                        group = distraction,
                        y=recall,
                        shape = distraction))+
  geom_line()+
  geom_point()+
  theme_classic()+
  labs(x = "Amount of Practice", y = "Recall")+
  scale_y_continuous(breaks = c(0,5,10,15,20), limits = c(0,20))+
  scale_x_continuous(breaks= c(2,4,8))+
  facet_wrap(~money_reward)

```
















