---
title: "Lab 7"
author: "Kristina Arevalo"
date: '`r format(Sys.time(), "%a %b %d %Y")`'
output: 
  pdf_document:
    latex_engine: xelatex
    toc: yes 
    toc_depth: 2
monofont: Times New Roman
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE, message = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(papaja)
library(tibble)
```

\newpage

# Problem 2
Complete the 2x2 factorial lab found here https://crumplab.github.io/statisticsLab/lab-10-factorial-anova.html, up to section 10.4.8. More specifically, your task is to follow that lab exercise to load in the data, transform the data into long-format, conduct a 2x2 between subjects ANOVA, and write a short results section reporting the main effects and interaction. (3 points)

## Load & Transform  data
```{r}

all_data <- read.csv("open_data/stroop_stand.csv")

RTs <- c(as.numeric(unlist(all_data[,1])),
         as.numeric(unlist(all_data[,2])),
         as.numeric(unlist(all_data[,3])),
         as.numeric(unlist(all_data[,4]))
         )

Congruency <- rep(rep(c("Congruent","Incongruent"),each=50),2)
Posture <- rep(c("Stand","Sit"),each=100)
Subject <- rep(1:50,4)

stroop_df <- data.frame(Subject,Congruency,Posture,RTs)
```

## ANOVA
```{r}
aov_out<- aov(RTs ~ Congruency*Posture, stroop_df) %>% summary()
aov_out
```
## Results

We conducted a 2x2 between subjects ANOVA where Congruency and Posture were our independent variables and the dependent variable was the mean reaction time to name the color. 
We found a main effect of Congruency, `r apa_print(aov_out)$full_result$Congruency `

There was also a main effect of Posture `r apa_print(aov_out)$full_result$Posture `

There was no two-way interaction between Congruency and Posture, F (1, 196) = .497, MSE = 13189.185, p < 0.481.

(I couldn't get the apa print function to print the interaction results for some reason it kept coming up as an error so I just wrote it out)

# Problem 4
In the conceptual section of this lab we used an R simulation to find the family-wise type I error rate for a simple factorial design with 2 independent variables. Use an R simulation to find the family-wise type I error rate for a factorial design with 3 independent variables. (3 points)

```{r}
save_sim <- tibble()

for(i in 1:1000){
n <- 12
factorial_data <- tibble(A = factor(rep(c("L1","L2"), each = n)),
                         B = factor(rep(rep(c("L1","L2"), each = n/2),2)),
                         C = factor(rep(c("L1","L2"), n)),
                         DV = rnorm(n*2,0,1))

output <- summary(aov(DV ~ A*B*C, data = factorial_data))

  sim_tibble <- tibble(p_vals = output[[1]]$`Pr(>F)`[1:7],
                       effect = c("A","B","C", "AxB","AxC", "BxC","AxBxC"),
                       sim = rep(i,7))

  save_sim <-rbind(save_sim,sim_tibble)
}

type_1 <- save_sim %>% 
  filter(p_vals < .05) %>% 
  group_by(sim) %>% 
  count()

dim(type_1)[1]/1000
```


